---
title: "Wine Quality Prediction"
output:
  pdf_document: default
  word_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Installation of the necessary packages (remove the # to install the packages):

```{r}
#install.packages("class", repos="http://cran.us.r-project.org")
#install.packages("gmodels", repos="http://cran.us.r-project.org")
library("class")
library("gmodels")
#install.packages("corrplot")
library("corrplot")
```

# Import the wine quality dataset.
```{r}

wine_quality<-read.csv2(file='C://Users/Doina/Desktop/winequality-white2.csv',
header = T, sep = ";", dec = ".",stringsAsFactors = FALSE)
```


# 1.Check the data characteristics. Is there missing data?


```{r}
str(wine_quality)
sum(is.na(wine_quality))

# There is no missing values
```



#2.What is the correlation between the attributes other than wine quality
```{r}
correlations<-cor(wine_quality[-c(12)])
correlations
corrplot(correlations, type = "lower")
# There is a very strong correlation between density and residual sugar,
# a strong correlation between total.sulfur.dioxide and free.sulfur.dioxide,
# a moderate correlation between density and total.sulfur.dioxide
# a strong negative correlation between alcohol and density
# a moderate negative correlation between fixed.acidity and ph
# a moderate negative correlation between residual.sugar and alcohol
# a moderate negative correlation between total.sulfur.dioxide and alcohol
# a moderate correlation between total.sulfur.dioxide and residual sugar

```


# 3.Graph the frequency distribution of wine quality.


```{r}
hist(wine_quality[,12], main="Wine quality distribution"  )
```
#4.Reduce the levels of rating for quality to three levels as high, medium and low.

```{r}
wine_quality$quality<-as.factor(wine_quality$quality)
wine_quality$quality<-factor(wine_quality$quality, levels = c("3",'4','5','6','7','8','9'),
labels = c("low","low","low","medium","medium","high", "high"))
# as the database describes the quality of the the wine according to 10 levels, 
#but in the database range between  3 and 9, I considered that levels: [3-5]="low",
#[6-7]="medium", [8-9]="high"
str(wine_quality)


```

# 5.Normalize the data set

```{r}
normalize <- function(x) {
return ((x - min(x)) / (max(x) - min(x))) }
wine_quality_n <- as.data.frame(lapply(wine_quality[-c(12)], normalize))
wine_quality_n <- cbind(wine_quality_n,wine_quality$quality)
head(wine_quality_n)

```

# 6.Divide the data to training and testinggroups


```{r}
set.seed(1)
index <- sample(1:nrow(wine_quality_n), 0.7 *nrow(wine_quality_n))
wine_quality_train <- wine_quality_n[index,]
wine_quality_test <- wine_quality_n[-index,]



```

# 7.Use the KNN algorithm to predict the quality of wine using its attributes.
```{r}
wine_quality_train_labels <- wine_quality_train[,12]
wine_quality_test_labels <- wine_quality_test[,12]

wine_quality_prediction<- knn(train = wine_quality_train[,1:11], test =wine_quality_test[,1:11],
cl = wine_quality_train[,12], k=7)
```

# 8.Evaluate the model performance 

```{r}
CrossTable(x=wine_quality_test_labels, y=wine_quality_prediction, prop.chisq=FALSE)

#The accuracy of the model =(TP+TN)/N= (297+757+0)/1470=71.7%. 
#We can observe that the model is not very good at predicting the high quality wines,
#but The model could be improved if the dataset whould be balanced. 
#To begin with the high quality has very few entries in the test set, so the algorithm
#is bias towards the majority, the medium quality wine.
```

